---
title: "Modifiers"
author: "Gina Reynolds, July 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "ninjutsu"]
    nature:
      highlightLines: yes
      ratio: '16:9'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
knitr::opts_chunk$set(fig.height = 6, out.width = "100%", comment = " ", cache = T, dpi = 300)
library(flipbookr)

```


```{r xaringan-themer, include = FALSE, eval = F}
# library(xaringanthemer)


# xaringanthemer::write_xaringan_theme(text_font_size = ".75cm",
#                                      text_color = "darkblue")

xaringanthemer::mono_accent(
  base_color = "#43418A",
  header_font_google = google_font("Montserrat"),
  text_font_google   = google_font("Montserrat", "200", "200i"),
  code_font_google   = google_font("Droid Mono"),
  text_font_size = ".7cm",
  code_font_size = ".7cm")
  

# write_xaringan_theme(
#   extra_css = list(
#     ".box-right" = list(
#       "height" = "300px",
#       "width" = "300px",
#       "position" = "absolute",
#       "top" = "33%",
#       "left" = "55%"
#     )
#   )
# )

```




---


<!-- # For teachers -->

<!-- This is a collection of thoughts about how to teach ggplot2, and any other grammar of graphics type data visualization system.   -->

<!-- It is tempting to rush to build a publication-ready plot right out of the gate.   -->

<!-- But maybe we should resist this temptation. Maybe we should "hang out" in each of the component zones for a bit.  Explore. Get comfy in a zone.  Then move on. -->


```{r, include=F}
load("data_products/manipulations.Rdata")
library(tidyverse)
```



---

# Modifiers

> *"The book is on the table"* 

--

or 

> *"The huge, paperback book is on the purple table"?*


---
class: center, middle

We are able to learn something about our data with very few inputs with ggplot2: 

--

declare data -- `ggplot(data = ?)`

--

declare aesthetic representation -- `aes(x = ?)`

--

declare geom  -- `geom_rug()`

--

... and we get a plot back.

---
class: center, middle

Part of the reason we have such a concise "grammar of graphics sentence" is there are convenient and reasonable defaults related to the aesthetics.  

--

In the first human, "The book is on the table", we also might a picture of the book and table -- it is some reasonable "default image" of books -- a genaric book, a genaric table.  

--

But we might use modifiers to change the picture, to specify that the table is not *brown* as we might imagine by default, but *purple*!

---
class: center, middle

The same is true in ggplot2, also, we can write short "sentences" and get a meaningful plot in return because of helpful defaults.  But we can use modifiers to change these defaults.  

---

# Modifiers


Modifiers adjust the default associated with aesthetic representations for:

--

1. Easy: the scale *labels* (the scale labels are the variable names by default, but they could be something else).  *Mastry is fairly easy. *

--

2. Medium: the *coordinate system* (the default coordinate system is cartesian (x horizontal, y vertical), but it could be something else - like polar).  *There's a bit more to know here.*

--

3. Hard: the aesthetic scales *themselves* (we have seen a "blues" scale by default for the continuous color aesthetic, but it could be greens, or yellows). *There's a lot to learn here!* 





---

name: labs

## Modifiers I: aesthetic scale labels

Aesthetic scales also get default *labels* which are the variable names.  We can also modify these.  

The syntax is labs(\* = "New label"), where the \* is the aesthetic name.  

`labs()` can also be used to add plot labels, plot title, caption, subtitle and tag.

Let's see how. 

---

```{r labs, eval = F, echo = F}
ggplot(data = gapminder_2002) +
  aes(x = gdpPercap) +
  aes(y = lifeExp) +
  geom_point() +
  aes(color = continent) +
  aes(shape = continent) +
  aes(size = gdpPercap) +
  aes(alpha = gdpPercap) + # this much is familiar #BREAK
  labs(x = "GDP per cap") + #BREAK
  labs(y = "Life Expectency") + #BREAK
  labs(color = "Continent") + #BREAK
  labs(shape = "Continent") + #BREAK
  labs(size = "GDP per cap") + #BREAK
  labs(alpha = "GDP per cap") + #BREAK
  # Note: labs is also used to adjust *plot* labels
  labs(title = "GDP per Cap versus\nLife Expectancy in 2002") + #BREAK
  labs(subtitle = "Data: Gapminder Data in R") + #BREAK
  labs(caption = "Visualization: @EvaMaeRey") + #BREAK
  labs(tag = "Plot 1") #BREAK
```


`r chunk_reveal("labs", break_type = "user")`

---

## Questions:

- When might you use labs(tag = )?
- What does the character string "\n" do in the place of a space in the plot title?
- Color and shape variables represent the same variable, and labs(color = \*) labs(shape = \*) have the same input, "Continent".  What does this mean for the legend? Change the labels to be different and see.
- Change the y label to an empty string, '""'.  What does this do?  



---

name: coords

## Modifiers II: x and y position coordinate system

A third default for how aesthetic mapping of *positions x and y* are manifest is the coordinate system.  The default is cartesian coordinates - where x lies on the horizontal and is pependicular to y.  

But let's quickly check out some of the different options for coordinates systems.  When we use a coord_* statement in ggplot, we can overwrite the default.    

---

`r chunk_reveal("coords", break_type = "rotate")`

```{r coords, eval = F, echo = F}
ggplot(data = gapminder_2002) +
  aes(x = gdpPercap) +
  aes(y = lifeExp) +
  geom_point() +
  coord_cartesian() + #ROTATE
  coord_equal() + #ROTATE
  coord_fixed(ratio = 400) + #ROTATE
  coord_flip() + #ROTATE
  coord_polar() + #ROTATE
```




---

name: scales

## Modifiers III: aesthetic mapping (representation) scales 

With the aes() statement we have declared a variable mapped to an aesthetic, like x position, color, line width, etc.  To go back and ask for a modified scale, we use scales\_\* statements.  These statements provide a quick way to update our preferences for all the scales used for mapped aesthetics.   

<!-- Start with a scales_x_* and look at the variety of scale options associated with the x position aesthetic.   -->

Let's look at how we modify the default scale with a scales\_\* statement.


---

### Scales statements


There are a variety of scaling alternatives for the x position aesthetic:  

- scale_x_log10()
- scale_x_reverse()
- scale_x_sqrt()


---


```{r x_scales, eval = F, echo = F}
ggplot(data = gapminder_2002) +
  aes(x = gdpPercap) +
  aes(y = lifeExp) +
  geom_point() +  # this much is familiar #BREAK
  scale_x_log10() + # overrides default  #BREAK
  scale_x_reverse() + # overrides log10  #BREAK
  scale_x_sqrt() # overrides reverse  #BREAK
```

`r chunk_reveal("x_scales", break_type = "user")`


---

### More aesthetic scale modifications

*Note that using multiple scales associated with a single aesthetic is not typical.  We are just doing it here to expose you quickly to a variety of possibilities.  ggplot2 uses whatever scale specification last declared in our examples.*

Analogously, we can use scale\_\*\_\* statements to modify other aesthetic scales.  Here are some examples

<!-- I think people find it easy to think about x and y position and adjusting those scales, and related adjustment like labs() and scale_*, but we teachers need to help extend these logics to the other aesthetics.  -->


- scale_x_log10() 
- scale_y_reverse(limits = c(100, 0)) 
- scale_color_viridis_d() 
- scale_shape_discrete(solid = F) 
- scale_size_continuous(breaks = c(100, 1000, 10000, 100000)) 
- scale_alpha(range = c(.5, 1)) # full transparency is 0


Let's tour several of those to get a taste.  

---

`r chunk_reveal("scales")`

```{r scales, eval = F, echo = F}
ggplot(data = gapminder_2002) +
  aes(x = gdpPercap) +
  scale_x_continuous(trans = "log10") +  
  aes(y = lifeExp) +
  scale_y_continuous(breaks = c(50,70)) +  
  geom_point() +
  aes(color = continent) +
  scale_color_viridis_d() +
  aes(size = gdpPercap) +
  scale_size_continuous(limits = c(0, 400000)) +
  aes(shape = continent) +
  scale_shape_discrete(solid = F) + 
  aes(alpha = gdpPercap) +
  scale_alpha(range = c(.4, .9))
```



---

## Questions:

<!-- to ponder -->

- What does the parameter "breaks" seem to control?

- What are some additional color scales that can be used?  Use auto complete in RStudio.  (i.e. type "scale_color" and then "tab" to see your options)

- What does "d" stand for in the scale_color_viridis_d()? Use the help pane.

- While you are in the help pane, see how would you set missing values (NA) to "black".

---

For a full tour of scale modifications check out the [scales second look](https://evamaerey.github.io/ggplot2_grammar_guide/scales). 






---

## Modifiers - Review

We have seen that there are three modifiers to aesthetic mapping.

- scale_*
- labs(* = )
- coords_? (applies x,y position aesthetics only)


To review, let's just look at modifiers for the horizontal *x* position aesthetic.

---

```{r modifiers_review, include = FALSE}
ggplot(data = gapminder_2002) +
  aes(x = gdpPercap) +
  aes(y = lifeExp) +
  geom_point() +
  labs(x = "GDP per cap") + # modify x label #BREAK
  coord_polar() + # modify x and y coordinate system 
  scale_x_log10()  # modify x scale
```


`r chunk_reveal("modifiers_review")`




```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 70%}
```


<!-- ```{r, dpi=400, eval = F} -->
<!-- gapminder %>%  -->
<!--   select(continent) %>%  -->
<!--   ggplot() + -->
<!--   aes(x = 1) + -->
<!--   aes(fill = continent) + -->
<!--   geom_bar(color = "white", size = .2) + -->
<!--   coord_polar(theta = "y") + -->
<!--   theme_void() + -->
<!--   scale_fill_viridis_d() + -->
<!--   theme(rect = element_rect(fill = "grey",  -->
<!--                             color = "grey",  -->
<!--                             linetype = "solid", -->
<!--                             size = 0)) -->

<!-- ggsave("pie.svg", dpi = 320,device = "svg") -->
<!-- ``` -->

<!-- --- -->

<!-- Or at least an organized twitter thread!?  Another idea for you: pull aes() out of ggplot(). Do you do it? Lot's of reasons to do it! Downside is currently most examples make use of nested approach. @grrrck does this too, and esquisse -->

<!-- ggplot(data = my_data) +  -->
<!--    aes(x = my_var) -->

<!-- Is it possible to do this with multiple geoms? I usually specify aes within the geom like  -->

<!-- ggplot(data) + -->
<!--   geom_point(aes(x, y)) + -->
<!--   geom_line(aes(x, y, group = id))  -->

<!-- I prefer this approach because it's explicit which aesthetics are bound to which geoms -->

<!-- My blog post, (which no one has probably ever read) exactly on this topic! https://evangelinereynolds.netlify.com/post/mapping-aesthetics/ … In general I'd say go global.  I think in general, most folks don't have a bunch of conflicts for aesthetics geom by geom (though occasionally yes?).  Let me know what you think! -->


<!-- To change the data used for a plot, use the %+% operator! Oh!!!  -->


<!-- # Stat_* -->

<!-- ## Univariate discrete -->



<!-- ```{r univariate_discrete, eval = F, echo = F} -->
<!-- ggplot(gapminder_2002) + -->
<!--   aes(x = continent) + -->
<!--   stat_count() + -->
<!--   geom_bar() # convenience geom -->
<!--             # default counting -->
<!-- ``` -->

<!-- --- -->

<!-- r chunk_reveal("univariate_discrete")` -->


<!-- --- -->

<!-- ```{r} -->
<!-- ggplot(data = gapminder_2002) + -->
<!--   aes(x = continent) + -->
<!--   aes(y = lifeExp) + -->
<!--   geom_point(alpha = .1) + -->
<!--   stat_summary( -->
<!--     fun.ymin = min, -->
<!--     fun.ymax = max, -->
<!--     fun.y = median -->
<!--   ) -->
<!-- ``` -->



<!-- --- -->

<!-- ```{r} -->
<!-- gapminder_2002 %>%  -->
<!--   mutate(seventy_plus = lifeExp > 60) %>%  -->
<!-- ggplot() + -->
<!--   aes(x = continent) + -->
<!--   aes(fill = seventy_plus) + -->
<!--   geom_bar(alpha = .2) -->
<!-- ``` -->

