---
title: "external themes"
subtitle: "a flipbook | made with Xaringan"
author: "<br><br>Gina Reynolds, April 2020"
date: "<br>"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["kunoichi", "ninjutsu"]
    nature:
      ratio: "16:9"
      highlightLines: true
---

```{r setup, echo = F, eval = T, message = F, warning = F }
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.height = 6, out.width = "100%", comment = " ", cache = T, message = F, warning = F, dpi = 300)

library(flipbookr)
```


---

# Base Plot

We'll work with the plot below to apply all the theme adjustments.  In the next slide, we'll build this plot and save it as the plot object `g`. 


---

`r chunk_reveal("the_plot")`


```{r the_plot, eval = T, echo = F}
library(tidyverse)
library(gapminder)
gapminder %>%
  filter(year == 1992) %>%
  ggplot() +
  aes(x = gdpPercap, y = lifeExp) +
  geom_point(alpha = .8, shape = 21, 
             fill = "white") +
  aes(col = continent, fill = continent) +
  geom_point(alpha = .3, shape = 21) +
  aes(size = pop) +
  scale_size(guide = F) +
  labs(title = "Wealth and expected longevity\n in 1992") +
  labs(subtitle = "Data from gapminder package in R") +
  labs(x = "GDP per Capita") +
  labs(y = "Life Expectency") +
  labs(tag = "Plot 1") +
  labs(col = "Continent", fill = "Continent") +
  labs(caption = "Vis: @EvaMaeRey with ggplot") ->
g
```


---

Now let's look at available themes from a couple external packages, {ggthemes} and {ggdark}. 

You will need to install these packages first in order to use them.  


---

`r chunk_reveal("ggthemes", break_type = "rotate")`

```{r ggthemes, eval = F, echo = F, out.width="60%"}
gapminder %>%
  filter(year == 1992) %>%
  ggplot() +
  aes(x = gdpPercap, y = lifeExp) +
  geom_point(alpha = .8, shape = 21, 
             fill = "white") +
  aes(col = continent, fill = continent) +
  geom_point(alpha = .3, shape = 21) +
  aes(size = pop) +
  scale_size(guide = F) +
  labs(title = "Wealth and expected longevity\n in 1992") +
  labs(subtitle = "Data from gapminder package in R") +
  labs(x = "GDP per Capita") +
  labs(y = "Life Expectency") +
  labs(tag = "Plot 1") +
  labs(col = "Continent", fill = "Continent") +
  labs(caption = "Vis: @EvaMaeRey with ggplot") + 
  ggthemes::theme_base() + #ROTATE
  ggthemes::theme_calc() +  #ROTATE
  ggthemes::theme_economist() + #ROTATE
  ggthemes::theme_economist_white() + #ROTATE
  ggthemes::theme_excel() + #ROTATE
  ggthemes::theme_excel_new() + #ROTATE
  ggthemes::theme_few() + #ROTATE
  ggthemes::theme_fivethirtyeight() + #ROTATE
  ggthemes::theme_gdocs() + #ROTATE
  ggthemes::theme_hc() + #ROTATE
  ggthemes::theme_igray() + #ROTATE
  ggthemes::theme_map() + #ROTATE
  ggthemes::theme_solarized() + #ROTATE
  ggthemes::theme_solarized_2() +  #ROTATE
  ggthemes::theme_solid() + #ROTATE
  ggthemes::theme_stata() + #ROTATE
  ggthemes::theme_tufte() #ROTATE
  ggthemes::theme_wsj() + #ROTATE
  ggdark::dark_theme_bw() + #ROTATE
  ggdark::dark_theme_classic() + #ROTATE
  ggdark::dark_theme_dark() + #ROTATE
  ggdark::dark_theme_gray() + #ROTATE
  ggdark::dark_theme_light() + #ROTATE
  ggdark::dark_theme_linedraw() + #ROTATE
  ggdark::dark_theme_minimal() + #ROTATE
  ggdark::dark_theme_test() #ROTATE
```



---

A different approach is that of {ggthemr}.  The idea is to set the theme and forget it.   It is not added to the ggplot, but set before plotting begins. 

---

`r chunk_reveal("ggthemr", break_type = "rotate")`

```{r ggthemr,include=F}
ggthemr::ggthemr('copper') #ROTATE
ggthemr::ggthemr('dust') #ROTATE
ggthemr::ggthemr('chalk') #ROTATE
ggthemr::ggthemr('flat') #ROTATE
ggthemr::ggthemr('flat dark') #ROTATE
ggthemr::ggthemr('camoflauge') #ROTATE
ggthemr::ggthemr('copper') #ROTATE
ggthemr::ggthemr('earth') #ROTATE
ggthemr::ggthemr('fresh') #ROTATE
ggthemr::ggthemr('grape') #ROTATE
ggthemr::ggthemr('grass') #ROTATE
ggthemr::ggthemr('greyscale') #ROTATE
ggthemr::ggthemr('light') #ROTATE
ggthemr::ggthemr('lilac') #ROTATE
ggthemr::ggthemr('pale') #ROTATE
ggthemr::ggthemr('sea') #ROTATE
ggthemr::ggthemr('sky') #ROTATE
ggthemr::ggthemr('solarized') #ROTATE
ggthemr::ggthemr_reset() #ROTATE
gapminder %>%
  filter(year == 1992) %>%
  ggplot() +
  aes(x = gdpPercap, y = lifeExp) +
  geom_point(alpha = .8, shape = 21, 
             fill = "white") +
  aes(col = continent, fill = continent) +
  geom_point(alpha = .3, shape = 21) +
  aes(size = pop) +
  scale_size(guide = F) +
  labs(title = "Wealth and expected longevity\n in 1992") +
  labs(subtitle = "Data from gapminder package in R") +
  labs(x = "GDP per Capita") +
  labs(y = "Life Expectency") +
  labs(tag = "Plot 1") +
  labs(col = "Continent", fill = "Continent") +
  labs(caption = "Vis: @EvaMaeRey with ggplot") 
```







```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 70%}

.column-left {
  float: left;
  width: 33.333%;
}

.column-right {
  float: right;
  width: 33.333%;
}

.column-center {
  display: inline-block;
  width: 33.333%;
}
```


