---
title: "Facets, quick themes, ggsave"
author: "Gina Reynolds, July 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "ninjutsu"]
    nature:
      highlightLines: yes
      ratio: '16:9'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
knitr::opts_chunk$set(fig.height = 6, out.width = "100%", comment = " ", cache = T, dpi = 300)
library(flipbookr)
```


```{r, include=FALSE}
library(tidyverse)
load("data_products/manipulations.Rdata")
```


---

name: facets

# Punctuation: Small multiples

Punctuation helps us break up language into digestible pieces. Breaking up visualization can also be helpful.  

Creating *small multiples* also known as *faceting* is a powerful way to make data more digestible.   In ggplot, small plots based on a categorical variable can be visualized side by side.  


```{r facets, eval = F, echo = F}
ggplot(data = gapminder_2002) +
  aes(x = gdpPercap / 1000) +
  aes(y = lifeExp) +
  geom_point() +
  aes(color = continent) +
  facet_wrap(~ continent) + #ROTATE
  facet_wrap(~ continent, ncol = 2) + #ROTATE
  facet_wrap(~ continent, scales = "free_x") + #ROTATE
  facet_wrap(~ continent, scales = "free_y") + #ROTATE
  facet_wrap(~ continent, scales = "free") + #ROTATE
  facet_null() + # overrides any facetting #ROTATE
  facet_grid(continent ~ pop > 10000000) + #ROTATE
  facet_grid(. ~ pop > 10000000) + #ROTATE
  facet_grid(continent ~ .) #ROTATE
```

---

`r chunk_reveal("facets", break_type = "rotate")`


---



---

name: themes

# Greetings: Themes

It may seem just cosmetic, but themes can have a big impact on the look and feel of you final plot.  The thematic elements are what first meet the eye.  For more information on fine adjustment of thematic elements, have a look at the [taming ggplot themes flipbook](https://evamaerey.github.io/little_flipbooks_library/taming_themes_in_ggplot/taming_ggplot_themes.html#1).  For now, check out how to change the look and feel of your graph with some preset themes.  

---

`r chunk_reveal("themes", break_type = "rotate")`


```{r themes, eval = F, echo= F}
ggplot(data = gapminder_2002) +
  aes(x = gdpPercap) +
  aes(y = lifeExp) +
  geom_point() +
  aes(color = continent) +
  theme_bw() + #ROTATE
  theme_classic() + #ROTATE
  theme_gray() + #ROTATE
  theme_light() + #ROTATE
  theme_minimal() + #ROTATE
  theme_void() + #ROTATE
  theme_light(base_size = 10, base_family = "Times") + #ROTATE
  theme_classic(base_family = "Rockwell") + #ROTATE
  theme_dark(base_line_size = 14) + #ROTATE
  theme_minimal(base_rect_size = 10) + #ROTATE
```



```{r, eval = F, echo = F}
(g <- ggplot(chic, aes(date, temp)) +
  geom_point())
old <- theme_set(theme_grey())
g
theme_update(
  panel.background = element_rect(
    fill = "tan",
    color = "black"
  )
) 
g
```



---

name: saving

# The written language: Save plots as objects and to disk

Now you should know how to build a ggplot.  How can you save that plot?  

First, you can save the entire plot as an object using the assignment operatore `<-`.  Below we call the plot `g`.  Then use the function ggsave() to save out your plot.  

```{r}
g <- ggplot(gapminder_2002) +
  aes(x = gdpPercap) +
  aes(y = lifeExp) +
  geom_point() 

ggsave(filename = "my_plot.png", 
       plot = g)
```





```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 60%}
```


<!-- ```{r, dpi=400, eval = F} -->
<!-- gapminder %>%  -->
<!--   select(continent) %>%  -->
<!--   ggplot() + -->
<!--   aes(x = 1) + -->
<!--   aes(fill = continent) + -->
<!--   geom_bar(color = "white", size = .2) + -->
<!--   coord_polar(theta = "y") + -->
<!--   theme_void() + -->
<!--   scale_fill_viridis_d() + -->
<!--   theme(rect = element_rect(fill = "grey",  -->
<!--                             color = "grey",  -->
<!--                             linetype = "solid", -->
<!--                             size = 0)) -->

<!-- ggsave("pie.svg", dpi = 320,device = "svg") -->
<!-- ``` -->

<!-- --- -->

<!-- Or at least an organized twitter thread!?  Another idea for you: pull aes() out of ggplot(). Do you do it? Lot's of reasons to do it! Downside is currently most examples make use of nested approach. @grrrck does this too, and esquisse -->

<!-- ggplot(data = my_data) +  -->
<!--    aes(x = my_var) -->

<!-- Is it possible to do this with multiple geoms? I usually specify aes within the geom like  -->

<!-- ggplot(data) + -->
<!--   geom_point(aes(x, y)) + -->
<!--   geom_line(aes(x, y, group = id))  -->

<!-- I prefer this approach because it's explicit which aesthetics are bound to which geoms -->

<!-- My blog post, (which no one has probably ever read) exactly on this topic! https://evangelinereynolds.netlify.com/post/mapping-aesthetics/ … In general I'd say go global.  I think in general, most folks don't have a bunch of conflicts for aesthetics geom by geom (though occasionally yes?).  Let me know what you think! -->


<!-- To change the data used for a plot, use the %+% operator! Oh!!!  -->


<!-- # Stat_* -->

<!-- ## Univariate discrete -->



<!-- ```{r univariate_discrete, eval = F, echo = F} -->
<!-- ggplot(gapminder_2002) + -->
<!--   aes(x = continent) + -->
<!--   stat_count() + -->
<!--   geom_bar() # convenience geom -->
<!--             # default counting -->
<!-- ``` -->

<!-- --- -->

<!-- r chunk_reveal("univariate_discrete")` -->


<!-- --- -->

<!-- ```{r} -->
<!-- ggplot(data = gapminder_2002) + -->
<!--   aes(x = continent) + -->
<!--   aes(y = lifeExp) + -->
<!--   geom_point(alpha = .1) + -->
<!--   stat_summary( -->
<!--     fun.ymin = min, -->
<!--     fun.ymax = max, -->
<!--     fun.y = median -->
<!--   ) -->
<!-- ``` -->



<!-- --- -->

<!-- ```{r} -->
<!-- gapminder_2002 %>%  -->
<!--   mutate(seventy_plus = lifeExp > 60) %>%  -->
<!-- ggplot() + -->
<!--   aes(x = continent) + -->
<!--   aes(fill = seventy_plus) + -->
<!--   geom_bar(alpha = .2) -->
<!-- ``` -->

