---
title: "Scales"
author: "Gina Reynolds, July 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "ninjutsu"]
    nature:
      highlightLines: yes
      ratio: '16:9'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
knitr::opts_chunk$set(fig.height = 6, out.width = "100%", comment = " ", cache = T, dpi = 300)
library(flipbookr)

```


```{r xaringan-themer, include = FALSE, eval = F}
# library(xaringanthemer)


# xaringanthemer::write_xaringan_theme(text_font_size = ".75cm",
#                                      text_color = "darkblue")

xaringanthemer::mono_accent(
  base_color = "#43418A",
  header_font_google = google_font("Montserrat"),
  text_font_google   = google_font("Montserrat", "200", "200i"),
  code_font_google   = google_font("Droid Mono"),
  text_font_size = ".7cm",
  code_font_size = ".7cm")
  

# write_xaringan_theme(
#   extra_css = list(
#     ".box-right" = list(
#       "height" = "300px",
#       "width" = "300px",
#       "position" = "absolute",
#       "top" = "33%",
#       "left" = "55%"
#     )
#   )
# )

```


```{r, include=F, cache=F}
load("data_products/manipulations.Rdata")
library(tidyverse)
library(gapminder)
```

---


<!-- # For teachers -->

<!-- This is a collection of thoughts about how to teach ggplot2, and any other grammar of graphics type data visualization system.   -->

<!-- It is tempting to rush to build a publication-ready plot right out of the gate.   -->

<!-- But maybe we should resist this temptation. Maybe we should "hang out" in each of the component zones for a bit.  Explore. Get comfy in a zone.  Then move on. -->


```{r, include=F}
load("data_products/manipulations.Rdata")
library(tidyverse)
```

---

## Each aesthetic (color, x-position, size), that does representation for us has a scale associated with it..

---

`r chunk_reveal("scales")`

```{r scales, eval = F, echo = F}
ggplot(data = gapminder_2002) +
  aes(x = gdpPercap) +
  scale_x_continuous(trans = "log10") +  
  aes(y = lifeExp) +
  scale_y_continuous(breaks = c(50,70)) +  
  geom_point() +
  aes(color = continent) +
  scale_color_viridis_d() +
  aes(size = gdpPercap) +
  scale_size_continuous(limits = c(0, 400000)) +
  aes(shape = continent) +
  scale_shape_discrete(solid = F) + 
  aes(alpha = gdpPercap) +
  scale_alpha(range = c(.4, .9))
```



---

# Concepts

- breaks - *values* to be labeled along aesthetic scales
- break labels - i.e. the exact text of the labels)
- *value* limits of variable representation
- transformations
- range of *aesthetic* values

---


`r chunk_reveal("scale_options", break_type = "rotate")`

```{r scale_options, include=F}
ggplot(data = gapminder_2002) +
  aes(x = gdpPercap) +
  aes(y = lifeExp) +
  aes(color = gdpPercap) +
  aes(alpha = lifeExp) +
  geom_point() +
  scale_x_continuous() + #ROTATE
  scale_x_continuous(breaks = c(1500, 15431)) + #ROTATE
  scale_y_continuous(breaks = c(50, 76, 81.3)) + #ROTATE
  scale_color_continuous(breaks = c(1500, 15431)) + #ROTATE
  scale_alpha_continuous(breaks = c(50, 76, 81.3)) + #ROTATE
  scale_x_continuous(breaks = c(1000, 1500), labels = c("1k", "15k")) + #ROTATE
  scale_y_continuous(labels = c("40y", "50y", "60y","70y", "80y")) + #ROTATE
  scale_color_continuous(breaks = c(1000, 1500), labels = c("1k", "15k")) + #ROTATE
  scale_alpha_continuous(labels = c("40y", "50y", "60y","70y", "80y")) + #ROTATE
  scale_x_continuous(limits = c(-1000,100000)) + #ROTATE
  scale_y_continuous(limits = c(30,100)) + #ROTATE
  scale_color_continuous(limits = c(-1000,100000)) + #ROTATE
  scale_alpha_continuous(limits = c(30,100)) + #ROTATE
  scale_x_continuous(trans = "log10") + #ROTATE
  scale_y_continuous(trans = "reverse") + #ROTATE
  scale_color_continuous(trans = "log10") + #ROTATE
  scale_alpha_continuous(trans = "reverse") + #ROTATE
  scale_x_continuous(expand = c(1,2)) + #ROTATE
  scale_y_continuous(expand = c(.1,.1)) + #ROTATE
  scale_color_continuous(expand = c(1,2)) + #ROTATE
  scale_alpha_continuous(expand = c(1,2)) #ROTATE
```





---


## Scales


There are a variety of scaling alternatives for the x position aesthetic:  

- scale_x_log10()
- scale_x_reverse()
- scale_x_sqrt()


---

`r chunk_reveal("x_trans", break_type = "rotate")`

```{r x_trans, include = F}
ggplot(data = gapminder_2002) +
  aes(x = gdpPercap) +
  aes(y = lifeExp) +
  geom_point() +
  scale_x_continuous(trans = "hms") + #ROTATE
  scale_x_continuous(trans = "sqrt") + #ROTATE
  scale_x_continuous(trans = "log10") + #ROTATE
  scale_x_continuous(trans = "log1p") + #ROTATE
  scale_x_continuous(trans = "log2") + #ROTATE
  scale_x_continuous(trans = "pseudo_log") + #ROTATE
  scale_x_continuous(trans = "reciprocal") + #ROTATE
  scale_x_continuous(trans = "reverse") + #ROTATE
  scale_x_continuous(trans = "sqrt") #ROTATE
```

---

`r chunk_reveal("x_break", break_type = "rotate")`

```{r x_break, include = F}
ggplot(data = gapminder_2002) +
  aes(x = gdpPercap) +
  aes(y = lifeExp) +
  geom_point() +
  scale_x_continuous(breaks = NULL) + #ROTATE
  scale_x_continuous(breaks = 1:3*10000) + #ROTATE
  scale_x_continuous(breaks = c(1000, 7090)) + #ROTATE
  scale_x_continuous(breaks = (1:10)^4) + #ROTATE
  scale_x_continuous(labels = scales::label_dollar()) + #ROTATE
  scale_x_continuous(labels = scales::label_percent()) + #ROTATE
  scale_x_continuous(labels = scales::label_pvalue()) + #ROTATE
  scale_x_continuous(labels = scales::label_comma()) + #ROTATE
  scale_x_continuous(labels = scales::label_math()) + #ROTATE
  scale_x_continuous(labels = scales::label_number_si(unit = "g")) + #ROTATE
  scale_x_continuous(labels = scales::label_number_si()) #ROTATE
```



---


```{r x_scales, eval = F, echo = F}
ggplot(data = gapminder_2002) +
  aes(x = gdpPercap) +
  aes(y = lifeExp) +
  geom_point() +  # this much is familiar #BREAK
  scale_x_log10() + # overrides default  #BREAK
  scale_x_reverse() + # overrides log10  #BREAK
  scale_x_sqrt() # overrides reverse  #BREAK
```

`r chunk_reveal("x_scales", break_type = "user")`


---

### More aesthetic scale modifications

*Note that using multiple scales associated with a single aesthetic is not typical.  We are just doing it here to expose you quickly to a variety of possibilities.  ggplot2 uses whatever scale specification last declared in our examples.*

Analogously, we can use scale\_\*\_\* statements to modify other aesthetic scales.  Here are some examples

<!-- I think people find it easy to think about x and y position and adjusting those scales, and related adjustment like labs() and scale_*, but we teachers need to help extend these logics to the other aesthetics.  -->


- scale_x_log10() 
- scale_y_reverse(limits = c(100, 0)) 
- scale_color_viridis_d() 
- scale_shape_discrete(solid = F) 
- scale_size_continuous(breaks = c(100, 1000, 10000, 100000)) 
- scale_alpha(range = c(.5, 1)) # full transparency is 0


Let's tour several of those to get a taste.  


---

## Questions:

<!-- to ponder -->

- What does the parameter "breaks" seem to control?

- What are some additional color scales that can be used?  Use auto complete in RStudio.  (i.e. type "scale_color" and then "tab" to see your options)

- What does "d" stand for in the scale_color_viridis_d()? Use the help pane.

- While you are in the help pane, see how would you set missing values (NA) to "black".


---
class: center, middle, inverse


# Color scales


---

`r chunk_reveal("viridis_c", break_type = "rotate")`

```{r viridis_c, include=F}
ggplot(data = gapminder_2002) +
  aes(x = gdpPercap) +
  aes(y = lifeExp) +
  geom_point(size = 4,
             alpha = .8) +
  aes(color = lifeExp) +
  scale_color_viridis_c(option = "cividis") + #ROTATE
  scale_color_viridis_c(option = "magma") + #ROTATE
  scale_color_viridis_c(option = "plasma") + #ROTATE
  scale_color_viridis_c(option = "inferno") + #ROTATE
  scale_color_viridis_c(option = "viridis") + #ROTATE
  scale_color_viridis_c(direction = -1) + #ROTATE
  scale_color_viridis_c(begin = .2, end = .5) + #ROTATE
  scale_color_viridis_c(breaks = 76, label = "76 years") + #ROTATE
  scale_color_viridis_c(guide = F) + #ROTATE
  scale_color_continuous(type = "gradient") + #ROTATE
  scale_color_continuous(type = "viridis") + #ROTATE 
  scale_color_gradient(low = "blue", high = "green") + #ROTATE 
  scale_color_gradient2(low = "blue", mid = "red", high = "green") #ROTATE 
```


---

`r chunk_reveal("viridis_d", break_type = "rotate")`

```{r viridis_d, include=F}
ggplot(data = gapminder_2002) +
  aes(x = gdpPercap/1000) +
  aes(y = lifeExp) +
  geom_point(size = 4,
             alpha = .8) +
  aes(color = continent) +
  facet_wrap(~continent) +
  scale_color_viridis_d(option = "viridis") + #ROTATE
  scale_color_viridis_d(option = "magma") + #ROTATE
  scale_color_viridis_d(option = "cividis") + #ROTATE
  scale_color_viridis_d(option = "plasma") + #ROTATE
  scale_color_viridis_d(option = "inferno") + #ROTATE
  scale_color_viridis_d() + #ROTATE
  scale_color_viridis_d(direction = -1) + #ROTATE
  scale_color_viridis_d(breaks = c("Asia", "Europe", "Oceania", "Africa", "Americas")) + #ROTATE
  scale_color_viridis_d(breaks = c("Asia")) + #ROTATE
  scale_color_viridis_d(breaks = c("Asia"), labels = c("A continent")) + #ROTATE
  scale_color_viridis_d(labels = c("AFRICA", "EUROPE", "OEANIA", "AFRICA", "AMERICAS")) + #ROTATE
  scale_color_viridis_d(guide = F) #ROTATE
```

---

```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 50%}
```


<!-- ```{r, dpi=400, eval = F} -->
<!-- gapminder %>%  -->
<!--   select(continent) %>%  -->
<!--   ggplot() + -->
<!--   aes(x = 1) + -->
<!--   aes(fill = continent) + -->
<!--   geom_bar(color = "white", size = .2) + -->
<!--   coord_polar(theta = "y") + -->
<!--   theme_void() + -->
<!--   scale_fill_viridis_d() + -->
<!--   theme(rect = element_rect(fill = "grey",  -->
<!--                             color = "grey",  -->
<!--                             linetype = "solid", -->
<!--                             size = 0)) -->

<!-- ggsave("pie.svg", dpi = 320,device = "svg") -->
<!-- ``` -->

<!-- --- -->

<!-- Or at least an organized twitter thread!?  Another idea for you: pull aes() out of ggplot(). Do you do it? Lot's of reasons to do it! Downside is currently most examples make use of nested approach. @grrrck does this too, and esquisse -->

<!-- ggplot(data = my_data) +  -->
<!--    aes(x = my_var) -->

<!-- Is it possible to do this with multiple geoms? I usually specify aes within the geom like  -->

<!-- ggplot(data) + -->
<!--   geom_point(aes(x, y)) + -->
<!--   geom_line(aes(x, y, group = id))  -->

<!-- I prefer this approach because it's explicit which aesthetics are bound to which geoms -->

<!-- My blog post, (which no one has probably ever read) exactly on this topic! https://evangelinereynolds.netlify.com/post/mapping-aesthetics/ … In general I'd say go global.  I think in general, most folks don't have a bunch of conflicts for aesthetics geom by geom (though occasionally yes?).  Let me know what you think! -->


<!-- To change the data used for a plot, use the %+% operator! Oh!!!  -->


<!-- # Stat_* -->

<!-- ## Univariate discrete -->



<!-- ```{r univariate_discrete, eval = F, echo = F} -->
<!-- ggplot(gapminder_2002) + -->
<!--   aes(x = continent) + -->
<!--   stat_count() + -->
<!--   geom_bar() # convenience geom -->
<!--             # default counting -->
<!-- ``` -->

<!-- --- -->

<!-- r chunk_reveal("univariate_discrete")` -->


<!-- --- -->

<!-- ```{r} -->
<!-- ggplot(data = gapminder_2002) + -->
<!--   aes(x = continent) + -->
<!--   aes(y = lifeExp) + -->
<!--   geom_point(alpha = .1) + -->
<!--   stat_summary( -->
<!--     fun.ymin = min, -->
<!--     fun.ymax = max, -->
<!--     fun.y = median -->
<!--   ) -->
<!-- ``` -->



<!-- --- -->

<!-- ```{r} -->
<!-- gapminder_2002 %>%  -->
<!--   mutate(seventy_plus = lifeExp > 60) %>%  -->
<!-- ggplot() + -->
<!--   aes(x = continent) + -->
<!--   aes(fill = seventy_plus) + -->
<!--   geom_bar(alpha = .2) -->
<!-- ``` -->

