---
title: "tuning ggplot themes"
subtitle: "a flipbook | made with Xaringan"
author: "<br><br>Gina Reynolds, May 2019"
date: "<br>"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: ["kunoichi", "ninjutsu"]
    nature:
      ratio: "16:9"
      highlightLines: true
---

```{r setup, echo = F, eval = T, message = F, warning = F }
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.height = 6, out.width = "100%", comment = " ", cache = F, message = F, warning = F, dpi = 300)

source("https://raw.githubusercontent.com/EvaMaeRey/little_flipbooks_library/master/xaringan_reveal_parentheses_balanced.R")
```


---

Some ideas from ["The Glamour of Graphics"](https://resources.rstudio.com/rstudio-conf-2020/the-glamour-of-graphics-william-chase)

- engage a wider audience
- aesthetic-usability effect
- respect for your audience

--

Glamour and Grammar connection

---

# Layout

- Alignment
  - Mostly left 
  - titles - *plot* left alignment
  - No head tilting
  - Get rid of labels if possible
  - Rethink legend position, or alternatives: labeling data, or color mapping in titles
  - Don't use border
  - Be gridline skeptical
  - Try to avoid legends
  - White space - use it!

---

# Typography

- Effect tone
- https://practicaltypography.com/
- https://fonts.google.com/ (source family)
- Typographic hiararchy - size, weight, spacing
- Don't use "oldstyle" numbers, do use *tabular* numbers

  
To be continued...  [](https://cran.r-project.org/web/packages/svglite/vignettes/fonts.html)
  

---

# Color

Check out the [color](color.html) slides. 

- also eyedropper tool
- use hue saturation value selection
- non-white backgrounds


---

# Table of Contents

- [Ugly Theme Choices](#uglyplot)

- Elements
  - [Text Elements](#textelements)
  - [Rectangular Elements](#rectangularelements)
  - [Line Elements](#lineelements)

- More
  - [Legends](#legend)
  - [Sundry](#sundry)
  - [Facets](#facet)

- [Preset Themes](#presetthemes)

---

# Visual Table of Contents

```{r, results='asis', echo=FALSE}
make_html_picture_link("figures/custom_theme.png", "#uglyplot")
make_html_picture_link("figures/text_components.png", "#textelements")
make_html_picture_link("figures/rect_components.png", "#rectangularelements")
make_html_picture_link("figures/line_components.png", "#lineelements")

```


```{r, results='asis', echo=FALSE}
make_html_picture_link("figures/legend.png", "#legend")
make_html_picture_link("figures/sundries.png", "#sundry")
make_html_picture_link("figures/small_multiples.png", "#facet")
make_html_picture_link("figures/theme_presets.png", "#presetthemes")
```



---

# Introduction

This is a focused extension of [the ggplot flipbook](https://evamaerey.github.io/ggplot_flipbook/ggplot_flipbook_xaringan.html).The aim is to illuminate theme statements. 




```{r, message = F, warning=F, eval = T}
options(scipen = 10) # adjust when scientific notation turns on
library(tidyverse)
library(gapminder)
```



---

# Base Plot

We'll work with the plot below to apply all the theme adjustments.  In the next slide, we'll build this plot and save it as the plot object `g`.  To this plot object we'll add the theme statements.  You might check out [The Tidyverse in Action](https://evamaerey.github.io/tidyverse_in_action/tidyverse_in_action.html) if you are puzzled about how to build `g`. The [ggplot2 Cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf) is also an excellent resource. 

```{r the_plot, eval = T, echo = F}
gapminder %>%
  filter(year == 1992) %>%
  ggplot() +
  aes(x = gdpPercap, y = lifeExp) +
  geom_point(alpha = .8, shape = 21, 
             fill = "white") +
  aes(col = continent, fill = continent) +
  geom_point(alpha = .3, shape = 21) +
  aes(size = pop) +
  scale_size(guide = F) +
  labs(title = "Wealth and expected longevity\n in 1992") +
  labs(subtitle = "Data from gapminder package in R") +
  labs(x = "GDP per Capita") +
  labs(y = "Life Expectency") +
  labs(tag = "Plot 1") +
  labs(col = "Continent", fill = "Continent") +
  labs(caption = "Vis: @EvaMaeRey with ggplot") ->
g
```



---

`r apply_reveal("the_plot")`





---

# Thematic adjustment overview

--

## Components

The components in a ggplot graph are usually things you can point to on a plot.  

For example the *legend box*, or the *y axis title*, or the *plot title*.   


--

## Elements

These components have a variety of elements.  For example the legend box has the perimeter of the box (the line) and it has the area of the box associated with it.  To "point" to specify the element parts of the components, functions like element_line(), element_rect(), and element_text() are used.  To delete an entire component, element_blank() can be used.

--

## Aesthetics

Aesthetics assignment for theme elements are similar to those available for geometric shapes in the plot. Aesthetics include color, size, line_type, and so on. 


---

# Graph Components

So, what components can be adjusted? (there are lots!)

> aspect.ratio

> axis.title, axis.title.x,
  axis.title.x.top, axis.title.x.bottom, axis.title.y, axis.title.y.left,
  axis.title.y.right, axis.text, axis.text.x, axis.text.x.top,
  axis.text.x.bottom, axis.text.y, axis.text.y.left, axis.text.y.right,
  axis.ticks, axis.ticks.x, axis.ticks.x.top, axis.ticks.x.bottom,
  axis.ticks.y, axis.ticks.y.left, axis.ticks.y.right, axis.ticks.length,
  axis.line, axis.line.x, axis.line.x.top, axis.line.x.bottom, axis.line.y,
  axis.line.y.left, axis.line.y.right
  
> legend.background, legend.margin,
  legend.spacing, legend.spacing.x, legend.spacing.y, legend.key,
  legend.key.size, legend.key.height, legend.key.width, legend.text,
  legend.text.align, legend.title, legend.title.align, legend.position,
  legend.direction, legend.justification, legend.box, legend.box.just,
  legend.box.margin, legend.box.background, legend.box.spacing
  
> panel.background, panel.border, panel.spacing, panel.spacing.x,
  panel.spacing.y, panel.grid, panel.grid.major, panel.grid.minor,
  panel.grid.major.x, panel.grid.major.y, panel.grid.minor.x,
  panel.grid.minor.y, panel.ontop 
  
> plot.background, plot.title,
  plot.subtitle, plot.caption, plot.tag, plot.tag.position, plot.margin
  
> strip.background, strip.background.x, strip.background.y,
  strip.placement, strip.text, strip.text.x, strip.text.y,
  strip.switch.pad.grid, strip.switch.pad.wrap



---
name: uglyplot

# An ugly plot, for locating components!

Now we'll create this plot to get a sense of how thematic modifications can be made, by 1) specifying the component to be adjusted 
2) the element of the component to be adjusted and 3) indicating the aesthetic adjustment.  

The adjustments made should make clear what components of the plot are called in ggplot. 

Later we'll walk through a more focused adjustments.  We'll go through time elements one at at time: text, rectangular, line, etc.  

```{r custom_theme, echo = F, eval = T, out.width="50%"}
g + 
  theme(rect = element_rect(fill = "snow3")) +
  theme(text = element_text(family = "Times New Roman")) +
  theme(axis.title = element_text(color = "slateblue4")) +
  theme(title = element_text(color = "seagreen")) +
  theme(axis.line = element_line(color = "orange")) +
  theme(legend.key = element_rect(fill = "violet")) +
  theme(axis.ticks = element_line(color = "blue")) +
  theme(axis.ticks.x = element_blank()) +
  theme(panel.background = element_rect(fill = "steelblue")) +
  theme(panel.grid = 
          element_line(size = 3, color = "thistle1")) +
  theme(legend.background = element_rect(fill = "goldenrod1")) +
  theme(legend.text = element_text(size = 7)) +
  theme(legend.position = "bottom") +
  theme(legend.margin = margin(t = 0, r = 20, 
                               b = 0, l = 40, unit = "pt")) +
  theme(aspect.ratio = .5) 
```

```{r, echo = F, warning=F, message=F, eval = T, fig.show='hide'}
get_what_save_what <- "custom_theme"
eval(parse(text = paste(knitr:::knit_code$get(get_what_save_what), collapse = "")))
ggsave(paste0("figures/", get_what_save_what, ".png"), dpi = 300)
```



---


`r apply_reveal("custom_theme")`





---

# What are the elements of these components?

Now we'll focus on different elements of the components: 

- element_line()
- element_rect()
- element_text()

Additionally, element_blank(), allows you to indicate that you don't want the component included in the plot at all. 


---


# Across-the-board adjustment

Across-the-board adjustment to all components with given elements can also be done using the *line*, *rect*, *text*, and *title* adjustments.
These will apply generally where no more specific adjustment has been made.  

> line

> rect

> text 

> title



---
name: textelements

# Focus on text elements!



```{r text_components, eval = T, echo = F, out.width="60%"}
g + 
  theme(plot.title.position = "plot") +
  theme(plot.tag = 
          element_text(color = "plum4", size = 25)) +
  theme(plot.title = 
          element_text(color = "plum4", 
                       size = 20,
                       lineheight = 2.5)) +
  theme(plot.subtitle = element_text(color = "plum4")) +
  theme(axis.title.y = 
          element_text(color = "plum4", 
                       size = rel(3))) +
  theme(axis.text.y = element_text(color = "plum4")) +
  theme(legend.title = 
          element_text(color = "plum4", size = 22)) +
  theme(legend.text = element_text(color = "plum4")) +
  theme(axis.title.x = element_text(color = "plum4")) +
  theme(axis.text.x = element_text(color = "plum4")) +
  theme(plot.caption = 
          element_text(color = "plum4", size = 20)) +
  theme(text = element_text(family = "Times")) + 
  theme(text = element_text(face = "bold.italic")) +
  theme(text = element_text(
    margin = margin(t = 3, r = 3, b = 3, l = 3, 
                    unit = "pt"))) +
  theme(text = element_text(angle = 2)) +
  theme(plot.title = element_text(hjust = .5)) +
  theme(axis.title.x = element_text(hjust = 1)) 
```


```{r, echo = F, warning=F, message=F, eval = T, fig.show='hide'}
get_what_save_what <- "text_components"
eval(parse(text = paste(knitr:::knit_code$get(get_what_save_what), collapse = "")))
ggsave(paste0("figures/", get_what_save_what, ".png"), dpi = 300)
```


---

`r apply_reveal("text_components")`


---

name: rectangularelements

# Focus on rectangular elements!


```{r rect_components, eval = T, echo = F, out.width="60%"}
g +
  theme(plot.background = 
          element_rect(fill = "purple", 
                       color = "darkolivegreen", 
                       size = 3)) +
  theme(panel.background = 
          element_rect(fill = "purple", 
                       color = "darkolivegreen",
                       size = 3)) +
  theme(legend.background = 
          element_rect(fill = "purple", 
                       color = "darkolivegreen", 
                       size = 3)) +
  theme(legend.key = 
          element_rect(fill = "purple", 
                       color = "darkolivegreen", 
                       size = 3)) +
  theme(panel.border = 
          element_rect(color = "magenta", 
                       fill = NA, size = 4))
```


```{r, echo = F, warning=F, message=F, eval = T, fig.show='hide'}
get_what_save_what <- "rect_components"
eval(parse(text = paste(knitr:::knit_code$get(get_what_save_what), collapse = "")))
ggsave(paste0("figures/", get_what_save_what, ".png"), dpi = 300)
```

---

`r apply_reveal("rect_components")`

---

name: lineelements

# Focus on line elements!


```{r line_components, eval = T, echo = F, out.width="60%"}
g +
  theme(axis.ticks.y = 
          element_line(size = 10, 
                       color = "blue")) +
  theme(axis.ticks.x = 
          element_line(size = 10, 
                       color = "blue")) +
  theme(panel.grid.major.y = 
          element_line(size = 2,
                       linetype = "dashed",
                       color = "blue",
                       lineend = "round")) +
  theme(panel.grid.major.x = 
          element_line(size = 2, 
                       arrow = arrow(),
                       color = "blue",
                       lineend = "round")) +
  theme(panel.grid.minor.y = 
          element_line(size = 1, 
                       linetype = "dotted",
                       color = "blue",
                       lineend = "butt")) +
  theme(panel.grid = 
          element_line(color = "blue")) +
  theme(axis.line.x.top = 
          element_line(size = 2,
                       color = "blue")) +
  theme(panel.ontop = TRUE,   
        panel.background = element_blank()) # in front of the data
```

```{r, echo = F, warning=F, message=F, eval = T, fig.show='hide'}
get_what_save_what <- "line_components"
eval(parse(text = paste(knitr:::knit_code$get(get_what_save_what), collapse = "")))
ggsave(paste0("figures/", get_what_save_what, ".png"), dpi = 300)
```

---

`r apply_reveal("line_components")`



---

# What aesthetics of these elements can be adjusted?

| aesthetic | adjustable for element of type: |
|-----|----|
| fill | *rect*  |
| color | *rect, line, text* |
| size |  *rect, line, text* |
| linetype | *rect, line* |
| lineend | *line* |
| arrow | *line* |
| family (font) | *text* |
| face | *text* |
| hjust | *text* |
| vjust | *text* |
| angle  | *text* |
| lineheight |  *text* |
| margin | *text* |
| debug | *text* |
| inherit.blank | *rect, line, text* |



---

name: legends

# More legend options!



```{r legend, eval = T, echo = F, out.width="60%"}
g +
  theme(legend.position = "bottom") +
  theme(legend.position = "none") +
  theme(legend.position = c(.8, .25)) +
  theme(legend.direction = "horizontal") +
  theme(legend.key = element_rect()) +
  guides(fill = guide_legend(size = 35))
```

```{r, echo = F, warning=F, message=F, eval = T, fig.show='hide'}
get_what_save_what <- "legend"
eval(parse(text = paste(knitr:::knit_code$get(get_what_save_what), collapse = "")))
ggsave(paste0("figures/", get_what_save_what, ".png"), dpi = 300)
```
---

`r apply_reveal("legend")`

---

name: sundries

# More modifications! (Sundries)


```{r sundries, eval = T, echo = F, out.width="60%"}
g +
  theme(plot.tag.position = "topright") + 
  scale_color_discrete(guide = FALSE) +
  scale_fill_discrete(guide = FALSE) +
  scale_x_continuous(position = "top") +
  scale_y_continuous(position = "right") +
  theme(aspect.ratio = .5) 
```

---


```{r, echo = F, warning=F, message=F, eval = T, fig.show='hide'}
get_what_save_what <- "sundries"
eval(parse(text = paste(knitr:::knit_code$get(get_what_save_what), collapse = "")))
ggsave(paste0("figures/", get_what_save_what, ".png"), dpi = 300)
```

`r apply_reveal("sundries")`


---

name: facet

# Focus on facet elements!


```{r small_multiples, eval = T, echo = F, out.width="60%"}
g + 
  facet_wrap(~ continent, 
             strip.position = "right") +
  theme(strip.background = 
          element_rect(fill = "pink")) +
  theme(strip.text.y = 
          element_text(color = "snow", size = 12)) +
  theme(strip.text.y = 
          element_text(angle = -80)) +
  theme(panel.spacing.x = 
          unit(1.5, "lines")) +
  theme(panel.spacing.y = 
          unit(2, "lines")) +
  theme(strip.switch.pad.grid = 
          unit(0, "cm"))
```

```{r getwhat, echo = F, warning=F, message=F, eval = T, fig.show='hide'}
get_what_save_what <- "small_multiples"
eval(parse(text = paste(knitr:::knit_code$get(get_what_save_what), collapse = "")))
ggsave(paste0("figures/", get_what_save_what, ".png"), dpi = 300)
```


---

`r apply_reveal("small_multiples")`

---

name: presetthemes

# preset themes

Next we'll have at different preset themes available in ggplot that can be applied in one line of code. 


```{r theme_presets, eval = T, echo = F, out.width="60%"}
g + 
  theme_bw() +
  theme_classic() +
  theme_gray() + 
  theme_light() +
  theme_linedraw() +
  theme_minimal() +
  theme_void() +
  theme_dark()
```


---

```{r, echo = F, warning=F, message=F, eval = T, fig.show='hide'}
get_what_save_what <- "theme_presets"
eval(parse(text = paste(knitr:::knit_code$get(get_what_save_what), collapse = "")))
ggsave(paste0("figures/", get_what_save_what, ".png"), dpi = 300) 
```

---

`r apply_reveal("theme_presets")`


---

# ggeasy

The theme() function might strike you as not easy to use, especially as you might not use it too frequently.  {ggeasy} contains a set of functions that let you do common tasks you could do with theme(), but may find tricky. 

---

`r flipbookr::chunk_reveal("easy", break_type = "rotate")`

```{r easy}
library(ggeasy)
g +
  easy_all_text_color(color = "green") + #ROTATE
  easy_all_text_size(size = 14) + #ROTATE
  easy_x_axis_labels_size(size = 20) + #ROTATE
  easy_legend_at(to = "bottom") + #ROTATE
  easy_adjust_legend(to = "right") + #ROTATE
  easy_plot_caption_size(size = 15) + #ROTATE
  easy_remove_y_axis() + #ROTATE
  easy_add_legend_title() + #ROTATE
  easy_adjust_legend() + #ROTATE
  easy_all_text_color() + #ROTATE
  easy_all_text_colour() + #ROTATE
  easy_all_text_size() + #ROTATE
  easy_change_legend() + #ROTATE
  easy_change_text() + #ROTATE
  easy_labs() + #ROTATE
  easy_legend_at() + #ROTATE
  easy_move_legend() + #ROTATE
  easy_plot_caption_size() + #ROTATE
  easy_plot_legend_size() + #ROTATE
  easy_plot_legend_title_size() + #ROTATE
  easy_plot_subtitle_size() + #ROTATE
  easy_plot_title_size() + #ROTATE
  easy_remove_axes()  + #ROTATE
  easy_remove_legend() + #ROTATE
  easy_remove_legend_title() + #ROTATE
  easy_remove_x_axis() + #ROTATE
  easy_remove_y_axis() + #ROTATE
  easy_rotate_labels() + #ROTATE
  easy_rotate_legend() + #ROTATE
  easy_rotate_legend() + #ROTATE
  easy_rotate_x_labels() + #ROTATE
  easy_rotate_y_labels() + #ROTATE
  easy_text_color() + #ROTATE
  easy_text_colour() + #ROTATE
  easy_text_size() + #ROTATE
  easy_x_axis_labels_size() + #ROTATE
  easy_x_axis_title_size() + #ROTATE
  easy_y_axis_labels_size() + #ROTATE
  easy_y_axis_title_size() #ROTATE
```



---










```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 70%}

.column-left {
  float: left;
  width: 33.333%;
}

.column-right {
  float: right;
  width: 33.333%;
}

.column-center {
  display: inline-block;
  width: 33.333%;
}
```


