---
title: "A Minimal #TidyTuesday Flipbook"
author: "Gina Reynolds, May 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["kunoichi", "ninjutsu"]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=F}
knitr::opts_chunk$set(fig.width = 6, message = F, warning = F, 
                      comment = "", cache = F)
```



```{r}
tibble(my_x = 1:12, my_y = rep(0:3, 3)) %>% 
  ggplot() +
  aes(x = my_x) +
  aes(y = my_y) +
  aes(col = my_y) +
  geom_point() +
  scale_color_viridis_c() +
  ggforce::geom_link2() 
```

```{r}
tibble(my_x = 1:12, my_y = rep(0:3, 3)) %>% 
  ggplot() +
  aes(x = my_x) +
  aes(y = my_y) +
  geom_point() +
  geom_rect(data = tibble(the_x = 1:3, the_y = 1:3),
            mapping = 
              aes(xmin = the_x,
                  ymin = the_y,
                  xmax = the_x + 1,
                  ymax = the_y + 1)
            )
  
```



And load the tidyverse.

```{r}
library(tidyverse)
```

And load the data from the tidytuesday github page. 

```{r}
download.file("https://github.com/rfordatascience/tidytuesday/raw/master/data/2019/2019-01-01/tidytuesday_tweets.rds", destfile = "temp.rds")
data_tidytuesday <- readRDS("temp.rds")
unlink("temp.rds")
library(tidyverse)
```




---

# Where we are going: 

We'll create this plot.  I have **echo** set to FALSE in the code chunk options here so that you don't see the code, and **eval** to TRUE so that the plot output is produced.  The code chunk is given the name "nobel", and this is used in the in apply_reveal function, which breaks down code by wrangling and plot statements. 

```{r link2, echo = F, eval = T, fig.height=5}
 data_tidytuesday %>% 
  mutate(day = lubridate::date(created_at)) %>% 
  group_by(day) %>% 
  tally() %>% 
  mutate(day_of_week = lubridate::wday(day, label = T)) %>% 
ggplot() + 
  aes(x = day) +
  # aes(xend = lag(day)) +
  aes(y = n) + 
  # aes(yend = lag(n)) +
  aes(color = n) +
  ggforce::geom_link2(n = 500) +
  geom_point(aes(alpha = day_of_week == "Tue")) +  
  geom_text(nudge_y = 1.5, label = "Tues.", size = 2.5,
            aes(alpha = day_of_week == "Tue")) +
  scale_alpha_discrete(range = c(0,1), guide = F) +
  theme_minimal() +
  scale_color_viridis_c( guide = F) +
  labs(title = "#TidyTuesday tweeted mostly Tuesdays") +
  labs(subtitle = "Fifteen hundred #TidyTuesday tweets from April to December 2018") +
  labs(caption = "@EvaMaeRey | Data: rtweets & #TidyTuesday | Per day calculations based on __ time zone") +
  labs(y = "Tweets per day") + 
  labs(x = "") +
  ggdark::dark_theme_light() 
```


`r apply_reveal("link2")`


```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 80%}
```


