---
title: "Aesthetic Mapping"
author: "Evangeline Reynolds"
date: "4/6/2020"
output: html_document
---

# Set-up... Data package and ggplot2 package

```{r setup}
library(gapminder)
library(ggplot2)
```


# Data manipulation

```{r data_manipulation}
gm_2002 <- gapminder[gapminder$year == 2002, ]
gm_2002_eu <- gm_2002[gm_2002$continent == "Europe", ]
continent_count <- data.frame(continent = c("Africa", "Americas", "Asia", "Europe", "Oceania"),
                              count = c(52, 25, 33, 30, 2))
```


## Task 1: First plot (same as Hans Rosling -- but a moment in time - 2002)

```{r}
ggplot(data = gm_2002) +
  aes(x = gdpPercap) +
  aes(y = lifeExp) +
  geom_point() +
  aes(size = pop) +
  aes(color = continent)
```


# Task 2: 

- Use *shape* instead of *color* to represent continent.

```{r}

```



# Task 3:

- Use *shape* and *color* to representing continent.

```{r}

```



# Task 4:

- use the aesthetic named *alpha* to represents gdp per cap


```{r}

```

What does this aesthetic seem to be?

> Your answer

---




# Task 5:

# plotting the gm_2002_eu data

create a ggplot where the data is gm_2002_eu (look at the result as you build your plots)

- have *x* represent the variable gdp per capita
- have *y* represent the variable life expectancy
- have *color* represent the variable  population
- use a **point** to take on these aesthetics (i.e add a geom_point() statement)
- *also* use a **rug** geometry to take on these aesthetics (i.e. add a geom_rug() statement)
- have *label* represent the variable continent
- use **text** *also* to take on these aesthetics (i.e. add a geom_text() statement -- will look a bit messy - don't worry about this).

```{r}

```

Answer the question:  What is the *default color scale* for continuous variables? 

> your answer here


---

# Task 6:


# The continent_count data


In a new code chunk create a plot:

  - where the data is continent_count (this much is set up for you)
  - where *x* represent the variable continent
  - where *y* represent the variable count
  - where the geom taking on the x and y position is **point**
  
```{r}
ggplot(data = continent_count)
```

---

# Task 7:


Now, using the code above, change geom_point() to geom_col().  *In addition* to *x* representing continent and *y* representing count:

- have *color* represent continent
- have *linetype* represent continent
- have *fill* represent continent
- have *alpha* represent count

**Execute the code as you build to see the changes you are triggering.**

```{r}

```

Based on your experience above, how is the aesthetic, *fill*, different from the aesthetic *color*?

> your answer here


---

# Task 8:


# geom_tile()

The next plot you'll create uses the **tile** geom, geom_tile().

Use the data, gm_americas, that is created below.  Then build the plot (it is started for you).

- have *x* position represent year
- have *y* position represent country
- have the geometric shape **tiles** take in the x and y aesthetics (i.e. add a geom_tile() statement)
- have *fill* represent life expectancy (lifeExp)

```{r}
gm_americas <- gapminder[gapminder$continent == "Americas",]
  
ggplot(data = gm_americas)
```



---

# Task 9:


# A scatter plot with cars

Examine the data "cars" that is available from "base" R. You can search cars in the help, if you need a reminder

Build a scatterplot with this data in ggplot2 (i.e. declare the data, declare desired aesthetic representation, and use geom_point() ).

```{r}
ggplot(data = cars)
```

---

# Task 10:


# A scatterplot with pressure

Examine the data "pressure" that is available from "base" R. You can search pressure in the help, if you need a reminder.

Build a scatterplot with this data in ggplot2 (i.e. declare the data, declare aesthetic representation, and use geom_point() ).

```{r}
ggplot(data = pressure)
```

---

# Task 11:


# geom_rect()

Execute the code below, then just answer the questions. 

The geom_rect() is a rectangle defined by xmin, xmax, ymin, and ymax.  Here were seeing a few new things.  First we can make adjustments on the fly to variables.  It might be best to show gdpPercap **in thousands**.  We can do that *on the fly*, by dividing by 1000.  Same with population --- writing all the zeros out is not a great choice, we can report in millions --- so we divide the variable through by 1000000. 

Can you do arithmetic operations with variables on the fly?

> yes or no?

We can also create something like a variable on the fly.  Where we see xmin set to zero, *this is like having a variable (a column in the data), that is filled from top to bottom with zeros*.  Same with ymin.  

We see xmin set to zero.  What is this like? You can retype the explanation imediately above, try to internalize of course.  

> your answer here...


We also see facet_wrap(~country) being used.  Describe the behavior of facet_wrap(~country) --- (i.e. What does it do?)

> your answer

What does the area of the rectangle capture?  Hint, the width is gdp per cap, and the height is population.  The area is multiplying these two giving you ...

> your idea here...


```{r}
options(scipen = 10)  # this tells r to not turn on scientific notation too soon

# this is a comment - if you want to write comments 
# use the hashtag at the beginning of the line
ggplot(data = gm_2002_eu) +
  aes(xmax = gdpPercap/1000) +
  aes(ymax = pop/1000000) +
  aes(xmin = 0) +
  aes(ymin = 0) +
  geom_rect() +
  facet_wrap(~country)
```

---

# Task 12:


Execute the code below.  

What aes() statements does the point use?  

> 

What aes() statements does the segement use?

>

You can also search geom_point and geom_segment in Rstudio help to answer these questions.  Scroll down in the "documentation" to the section "Aesthetics".


```{r}
options(scipen = 10)
ggplot(data = gm_2002_eu) +
  aes(x = gdpPercap/1000) +
  aes(y = pop/1000000) +
  aes(xend = 0) +
  aes(yend = 0) +
  geom_segment() +
  geom_point() +
  facet_wrap(~country)
```


---

# Task 13: Deconstructing a plot.


Examine code below, and execute.  

What is the name of the object which contains the dataset for the plot?  

> your answer

What variable is represented by the x position?

> your answer

What variable is represented by the y position?

> your answer

What is the geometric shape that is taking on the aesthetic representations?

> your answer


Is there a probelm with this visual representation?  If yes, describe that problem...

> your answer


```{r}
ggplot(data = diamonds) +
  aes(x = depth) +
  aes(y = price) +
  geom_point()
```

# Task 14: Write the concise "grammar of graphics" definition of a data visualization/statistical graphic (choose either the colloquial or academic).


> A data visualization is ...

or 

> A statistical graphic is ...

Note the parallel with construction of ggplots:  

declare data + 

declare aesthetic representations + 

declare geometry that will take on aesthetics.


