---
title: 'visualizing distributions lab'
author: "Evangeline Reynolds"
date: "2/2/2020"
output: html_document
---


For this assignment, consult the [ggplot2 Cheat Sheet]( https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf) from RStudio and [geoms: a closer look](https://evamaerey.github.io/ggplot2_grammar_guide/geoms.html#1).  

On the ggplot2 Cheat Sheet, there are panels that talk about relevant to  "univariate" description (uni = one) and "bivariate" description (bi = two).  What exactly are the titles of these panels on the cheat sheet?

> your answer

Within these panels, what are some of the subcategories? 


```{r setup, include = F}
knitr::opts_chunk$set(message = F)
library(tidyverse)
library(gapminder)
```


```{r}
gapminder$long_life <- gapminder$lifeExp >= 70 
gapminder$life_cat <- ifelse(gapminder$long_life, "70+", "<70")
gapminder_2007 <- gapminder[gapminder$year == 2007,]
```

# Plotting

With the `gapminder_2007` object you've created, create the following visualizations and state the type of variable is each variable is (either **continuous** or **discrete**?)

## Visualizing bivariate (2 variables joint) distributions

### 1. visualize the relationship between gdp per capita and population

> variable types (**continuous** or **discrete**?): 

Use **two** geometric layers (geom_*) in the same plot.  Remember, the second geom in the code will plot atop the first geom; make sure both are somewhat visible.  

```{r}

```

### 2. visualize the relationship between gdp per capita and continent

> variable types (**continuous** or **discrete**?): 

Use **two** geometric layers in the same plot.


```{r}

```

###  3. visualize the relationship between continent and "life_cat"

> variable types (**continuous** or **discrete**?): 

Use **one** geometric layers in the same plot; consult the RStudio cheat sheet.  (you may also add geom_count, but this is bonus.)

```{r}

```


## Univariate (single variables) distribution


### 1. Distribution of all countries' population

Use **two** geometric layers in the same plot.  

```{r}

```

### 2. Distribution within life span categories (life_cat).  (The number of countries within the categories)

> variable type (**continuous** or **discrete**?): 

Below, the plot have already been created for you, in 2 ways.  Both versions of the code yeild the same plot.  Which geometric object geom_col or geom_bar has included an intermediate step of data manipulation done automatically for you? 

>

Try to add a geom_point() layer to each plot.  In which is this possible?  Look at the help pane for geom_point().  Try to explain the behavior. 

>

```{r}
gapminder_2007 %>% 
  count(life_cat) %>% 
  ggplot() +
  aes(x = life_cat) +
  aes(y = n) +
  geom_col()
```

```{r}
gapminder_2007 %>% 
  ggplot() +
  aes(x = life_cat) +
  geom_bar()
```


---

# On your own

Check out the short description of data, and look at the objects created in the following code chunk `game_goals`, `office_ratings`, `sf_trees`, or `tb_age`.

- [SF trees](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-01-28/readme.md)
- [nhl](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-03-03/readme.md)
- [The Office](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-03-17/readme.md)
- [Brain](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-03-24/readme.md)

```{r}
# nhl data
game_goals <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-03/game_goals.csv')

# the office series ratings
office_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-17/office_ratings.csv')

sf_trees <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-28/sf_trees.csv')

# brain
tbi_age <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-24/tbi_age.csv')
```

Build one or two plots with one of these data sets...

```{r}
# ggplot(data = ?) # uncomment and begin
```

