---
title: "Flipbooks"
author: "Gina Reynolds, November 2019"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: ["kunoichi", "ninjutsu"]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=F}
knitr::opts_chunk$set(fig.width = 6, message = F, warning = F, 
                      comment = "", cache = T)
```


# What are Flipbooks?
# Show an one example - ggplot2 flipbook
# What are their advantages (Don Norman, the Design of Everyday Things)
 - natural mapping (matching code-output temporal changes)
 - affordances (ease of use - motion detection, name, automatic parsing - low additional burden on creator)
 - feedback (minimal changes mean that feedback is digestible)
# How is it implemented?
  # automation... (technical side)
  # And how YOU can make one ...
# Extending the basic
  ## adding connectors, allowing code wrapping, using a parser (making it safe for human consumption)
  ## using %>% .[] - data.table
  ## Regular Assignment
  ## Non sequential reveal (not yet implemented)
  ## Traditional presentations
  ## Ensembles (not yet stable)
  ## python, C++ (not yet implemented)
  ## Visual Table of contents - "You do not need a slide for the Table of Contents!"
  ## Flipbook mini (not yet final)
# resources
  ## package
  ## flipbooks library (my uses)
  ## minimal examples
  ## video?
# caveats
  # irreproducibility across platforms (fonts)
  # output not supported by remark.js, not going to work
  # may load slow (or not at all - leaflet)
# future...???

---

# Introduction

This is a minimal example to demonstrate how to create a flipbook with data from #TidyTuesday.  It walks through data wrangling and plots pipelines made with the Tidyverse.  The functions that make this possible are the work of Emi Tanaka, Garrick Aden-Buie and myself, and are built for Xaringan, an Rmarkdown file type for creating presentation slides; the functions make use of the function `knitr:::knit_code$get()`.

The code to create the flipbook is an .Rmd that you can download [**here**](https://raw.githubusercontent.com/EvaMaeRey/little_flipbooks_library/master/tidytuesday_minimal_example/tidytuesday_minimal_example.Rmd).

---

Interested in more flipbooks? Check out 

- [the ggplot flipbook](https://evamaerey.github.io/ggplot_flipbook/ggplot_flipbook_xaringan.html)
- [The Tidyverse in Action](https://evamaerey.github.io/tidyverse_in_action/tidyverse_in_action.html)
- [the ggplot2 grammar guide](https://evamaerey.github.io/ggplot2_grammar_guide/ggplot2_grammar_guide)

and 

- [the little flipbooks libraray](https://raw.githubusercontent.com/EvaMaeRey/little_flipbooks_library)


For more about Xaringan:

- [Xaringan presentation slides](https://slides.yihui.name/xaringan/)

The sequential workflow of the Tidyverse makes incremental display of pipelines and ggplot statements ideal:  

- [www.tidyverse.org](https://www.tidyverse.org/)


---


# What's the slow ggplot style?

"Slow ggplot" just means working more incrementally than is typical. Elements of the approach are as follows:

- pulling out aes() from the ggplot() function:  
- using fewer functions; example - using labs() to add a title instead of ggtitle()
- using functions multiple times; example aes(x = var1) + aes(y = var2) rather than aes(x = var1, y = var2)
- using base R functions and tidyverse functions.  For other packages, the :: style to call them
- write out arguments (no shortcuts)  aes(x = gdppercap) not aes(gdppercap)
- order ggplot commands so that reactivity is obvious; scale adjustments to aesthetics might also be near the aesthetic declaration. 


---

Here, I contrast the usual plotting method to slow ggplotting:

Usual approach:

```{r, eval = F}
ggplot(my_data, aes(var1, y = var2, col = var3)) +
  geom_point() +
  ggtitle("My Title") +
  labs(x = "the x label", y = "the y label", col = "legend title")
```

Using slow ggplotting: 

```{r, eval = F}
ggplot(data = my_data) +
  aes(x = var1) +
  aes(y = var2) +
  geom_point() +
  aes(col = var3) +
  labs(y = "the y label") +
  labs(x = "the x label") +
  labs(col = "legend title") +
  labs(title = "My title")
```




---

# Set up

Okay. Let's load the the `flipbookr` and the `tidyverse` packages.

```{r, cache = F}
library(flipbookr)
library(tidyverse)
```

And load the data from the tidytuesday github page. 

```{r}
# nobel_winners <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-05-14/nobel_winners.csv")
# source("../../../../../R/flipbook_mini.R")
# source("../../../../../R/base_parse_reveal_xaringan.R")
# source("../../../../../R/non_sequential_reveal.R")
```

---

# Where we are going: 

We'll create this plot.  I have **echo** set to FALSE in the code chunk options here so that you don't see the code, and **eval** to TRUE so that the plot output is produced.  The code chunk is given the name "nobel", and this is used in the in apply_reveal function, which breaks down code by wrangling and plot statements. 

```{r , echo = F, eval = T, fig.height=5}
g <-  cars %>%             # the data  #REVEAL
  filter(speed > 4) %>%  # subset
 ggplot() +              # pipe to ggplot
  aes(x = speed) +
  aes(y = dist) +
  # Describing what follows
  geom_point(alpha = .3) + #REVEAL
  aes(color =
  speed > 14
  ) %+%
  cars

str(g)


g1 <- ggplot(cars, aes(x = speed)) 
g2 <- ggplot(cars, aes(x = speed))


for (i in 1:9) {
  
identical(g2[[i]], g1[[i]]) %>% 
    print()
  
}

g2$mapping
g1$mapping
g1$labels


```
 

 
---

# How do we get there?

In the next slide, we'll walk through the code that produces this plot, and the output along the way. We use the code `apply_reveal("nobel")` in-line to access the code from the code chunk called *nobel*.
 
---

`r apply_reveal("nobel")`




```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 80%}
```
